---
title: "Complete Palette Gallery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Complete Palette Gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 3
)
library(wixarikaRt)
```

## Sequential Palettes

### Tatewari Warm
```{r tatewari-warm}
wixarika_show_pal("tatewari_warm")
```

Inspired by Grandfather Fire. Perfect for warm data progressions.

### Kauyumari Forest & Sky
```{r kauyumari}
wixarika_show_pal("kauyumari_forestsky")
```

Sacred deer spirit colors, blending forest and sky.

### Hikuri Sequential
```{r hikuri-seq}
wixarika_show_pal("hikuri_seq")
```

Smooth gradient inspired by sacred peyote.

## Diverging Palettes

### Nierika Diverging
```{r nierika}
wixarika_show_pal("nierika_diverging")
```

Portal between worlds - cool to warm.

### Haramara Tatewari
```{r haramara}
wixarika_show_pal("haramara_tatewari")
```

Ocean to fire - teal through cream to gold.

## Qualitative Palettes

### Muvieri Vivid
```{r muvieri}
wixarika_show_pal("muvieri_vivid")
```

Ceremonial arrows - bold, contrasting colors.

### Mapa de Lenguas
```{r mapa}
wixarika_show_pal("mapa_lenguas")
```

Based on the historic indigenous languages map of Mexico.

## Colorblind-Friendly Options

### Tsikuri Qualitative (CB)
```{r cb-qual}
wixarika_show_pal("tsikuri_q_cb")
```

### Nakawe Sequential (CB)
```{r cb-seq}
wixarika_show_pal("nakawe_seq_cb")
```

### Tsikuri Diverging (CB)
```{r cb-div}
wixarika_show_pal("tsikuri_div_cb")
```
